noam_data_exploration_v1

### 2024-07-17 noam_data_exploration_v3.Rmd
Added another possible glmm approach to analyzing adiv that's more analagous to what other studies have done

@@

alpha_diversity$Cover <- rowSums(wide_transect_level[5:74])/100

diversity_stability <- alpha_diversity %>%
  dplyr::group_by(Site, Habitat, Site_habitat, Transect) %>%
  dplyr::summarise(Richness_mean = mean(Richness), 
            Richness_se = sd(Richness)/sqrt(n()),
            Richness_cv = CV(Richness),
            Shannon_mean = mean(Shannon), 
            Shannon_se = sd(Shannon)/sqrt(n()),
            Shannon_cv = CV(Shannon),
            Cover_mean = mean(Cover),
            Cover_se = sd(Cover)/sqrt(n()),
            Cover_cv = CV(Cover))

@@

>>>>

# probably the closest to what others have done:
richness_mod_2 <- glmmTMB(
  Richness ~ Year_factor*Habitat + (1 | Site/Transect/Location) + ar1(Year_factor + 0 | Location),
  family = poisson,
  data = alpha_diversity_quad,
  control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000))
) # overspecified but going with it for now
summary(richness_mod_2) # looks like strong autocorrelation, which makes sense to me
Anova(richness_mod_2)
plot(allEffects(richness_mod_2))
emmeans(richness_mod_2, pairwise ~ Habitat | Year_factor) # if we want to compare habitats year to year

@@

alpha_diversity$Cover <- rowSums(wide_transect_level[ , -which(names(wide_transect_level) %in% meta_cols)])/100
alpha_diversity$Stratum <- paste(alpha_diversity$Site, "-", alpha_diversity$Habitat, "-", alpha_diversity$Transect)
# stability_transect <- community_stability(df = transect_level,
#                                           time.var = "Year",
#                                           abundance.var = "Mean_cover",
#                                           replicate.var = "Stratum") # using codyn measure of stability
# NOTE: codyn community_stability() is just the inverse of CV
diversity_stability <- alpha_diversity %>%
  dplyr::group_by(Site, Habitat, Site_habitat, Transect, Stratum) %>%
  dplyr::summarise(Richness_mean = mean(Richness), 
            Richness_se = sd(Richness)/sqrt(n()),
            Richness_cv = CV(Richness),
            Shannon_mean = mean(Shannon), 
            Shannon_se = sd(Shannon)/sqrt(n()),
            Shannon_cv = CV(Shannon),
            Shannon_stability = 1/CV(Shannon), # Stability as defined by codyn and many others (including Pete in 2022 paper)
            Cover_mean = mean(Cover),
            Cover_se = sd(Cover)/sqrt(n()),
            Cover_cv = CV(Cover),
            Cover_stability = 1/CV(Cover)) # as above

library(ecofolio) # possible other method for examining portfolio effects

@@

```{r diversity stability quadrat level}
